One Amazon Problem? Problem-constructions and Transnationalism in Brazilian Presidential Discourse since 1985

### 2023-02-16 Submission/Re-submission/draft/draft_resubmission_shorten.rmd
all new changes

---
title: How has the Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Sociology and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences and affiliated to the Centre for International Environmental Studies.]
date: 13 February 2023
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\textbf{Word Count:}#1}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: The Amazon is often approached as a transnational environmental problem without a discussion of how key actors represent it differently in space and time. We combine theories of problem-representation and transnationalism to analyze where, when, and how the Brazilian Amazon has been constructed as a problem. We develop a framework that explains variation in transnational problem-constructions based on speech setting and problem urgency To test our framework, we use supervised machine learning to classify statements about the Amazon in 6240 presidential speeches since 1985. We find that when presidents speak outside (or inside) the Amazon and environmental problems are less (or more) urgent, presidents are likelier to highlight issues of environmental conservation (or economic integration). Furthermore, we show how we are moving from a period when the Amazon as a binary of either environmental or an economic issue has been overcome, to a moment when diverse problem-constructions are deployed.

---

```{=latex}
\keywords{ discourse analysis, transnational governance, environmental policy, Brazilian Amazon, supervised learning, deforestation}
```

```{=latex}
\wordcount{ 8190 (including abstract, text, references, and footnotes)}
```

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(ggplot2)
library(plm)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

> *"[...] I will fight against the NGOs on this issue of associating the increase in deforestation to our agricultural expansion. First of all, these NGOs should go plant trees in their own countries [...]" (President Lula, Cuiaba - MT, 2007).*

# 1. Introduction

*We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem.* Each of these statements represent the Amazon as a particular problem: national sovereignty, economic integration, social development, and environmental conservation, respectively. These specific problems touch on common understandings of the Amazon as part of the larger socio-cultural history of the country. Brazilian governments have been portrayed in the literature as having held differing views on the Amazon, with some prioritizing national sovereignty and economic integration (military dictatorship and Bolsonaro), while others prioritize environmental conservation (Sarney and Lula). However, this oversimplifies government policies and discourse, which in reality are more intricate and contradictory.

By combining @bacchi1999’s theory of problem representation and @keck1998's theory of transnationalism, we place the way the objects of policies are constructed in discourse at the center of our analysis. We offer a framework that explains variation in discursive problem-constructions depending on speech setting and problem urgency as actors participate and shape transnational politics. We focus on speeches, rather than policies, because they allow us to capture the diversity of ways the Amazon, as a policy object that demands dedicated attention, has been discursively constructed as a problem. While many analyses of Brazilian environmental discourse and policies intend to dissect its national and international nature, they rarely adopt methodological strategies that distinguish how important actors behave differently in both scenarios [@hochstetler2007]. Our framework and methodology help identify and account for these variations. We show, for instance, that when speaking to farmers in Cuiaba, one of Brazil’s agricultural powerhouses, in a moment when the environmental problem is urgent, it is likely that any president would sidestep environmental issues, as President Lula did in the opening epigraph.

To answer how the Amazon has been constructed as a problem, we create a dataset containing 6240 official speeches by all Brazilian presidents since 1985. Next, we identify all instances where the presidents mention the Amazon. Using an inductive approach, we develop a codebook to categorize how each statement constructs the Amazon as one or multiple problems related to national sovereignty, economic integration, social development, or environmental conservation. Then, we manually code a random subset of the statements and use supervised machine learning to label the remaining. Finally, we conduct a descriptive and inferential analysis of the coded data.

We have four main findings. First, presidents adopt different constructions of the Amazon in different settings according to how they expect to leverage audiences for symbolic and/or  material gains. Second, presidents often focus on alternative problem-constructions when problems become more urgent; whereas they boast about positive outcomes when problems are less urgent. Third, high-profile environmental related events, which receive strong international media attention, increase the importance of the Amazon in presidential speeches. Alternatively, relevant Amazonian-related policies appear to not to attract much attention. Finally, while all presidents deploy multiple problem-constructions across space and over time, we are now in a period when the dictum between understanding the Amazon as an environment or as an economic issue has been overcome, at least discursively. Conceptually, our research shows how focusing on problem-constructions highlight what certain objects of policy are assumed to be, and when they are ignored or privileged, based on the transnational setting of the speech and the urgency of aligned problems.

We begin our article by outlining the connection between problem-constructions, transnationalism, and presidential speeches. Next, we present our research design including data description and methodological approach. We open our analysis with our inferential analysis and move to explain developments over time. We conclude by summarizing the empirical and conceptual contributions of the article as well as providing avenues for future research.

@@

Social scientists often describe federal governments as proponents of a cohesive set of policies toward the Amazon [see @drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. The 1964 military dictatorship, for example, is associated with securing sovereignty in the region by populating it and integrating it into the national economy [@drummond2006]. Whereas, governments from the late-1980s up to the late-2000s are associated with a turn towards policies that focus on environmental conservation of the Amazon [@padua2012]. The presidencies of Rousseff (2011-2016) and Temer (2016-2018) are connected to the de-prioritizing of environmental conservation policies, while Bolsonaro (2019-2022) with the dismantling of environmental policies [@capobianco2019]. Although these works are important to understand how different governments acted towards the Amazon, they represent governments monolithically by bundling them into cohesive policy periods, or policy cycles, masking government diversity in policies and discourse.

We argue that the reason behind these monolithic representations of governments in literature about the Amazon is conceptual: many studies of the region contain an explicit or implicit assumption of what the “Amazon problem” is. There is more variation within a specific government, though, than suggested by the policy cycle approach. Social scientists have shown how governments have contradictory policies [@lopez2023; @fairfield2017; @poncedeleon2021]. The Amazon has often been seen as a key biome in global climate-mitigation, and consequently, as an environmental problem. If the Amazon is thought of as an “environmental” problem, then it demands an “environmental” solution. However, to build a more comprehensive account of the Amazon as forest, region, and peoples we need a framework that assumes the possibility of varied problem-constructions beyond Amazonian environmental nature.

@bacchi1999 [p. 2] places the question of “what is the problem represented to be” in the center of policy studies. These objects of policies have a cultural dimension as they take “shape within specific historical and national or international contexts” [@bacchi2009, p. 10]. The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed. Hence, discourse represents objects of policies as instances of a specific problem, which can be explicit or implicit. Problem-representation matters because it carries implications related to who is involved both in the root causes and in the solutions of the problem [@bacchi1999].

@@

Presidential discourses are a good instance to identify problem-constructions over time. While the relationship between presidential speeches and policy may not be linearly causal, speeches have the power to introduce, define, and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004]. This is particularly true in the Brazilian context where the executive branch of government, led by the president, has historically been the primary player in shaping public policy, especially after the 1988 constitution[@macaulay2017]. In Brazil’s hybrid system, often referred to Coalitional Presidencialism [@couto2021], presidents build legislative support for a broad agenda through the distribution of positions in the state bureaucracies. This means that presidents have enough power and support to implement the agenda they deem appropriate. We thus consider presidential speeches an adequate instance to identify various problem-constructions for the same policy object.

@@

The possibility of different problem-constructions for the same policy object in political discourse, requires an explanation of why they vary. There are two notable contenders for tying the problem-construction to international politics: Putnam’s two-level game and Keck and Sikkink’s transnational networks.

@putnam1988 argues that the outcomes of international negotiations lie within the overlap between the interests (and pressure) of domestic and international groups. At the national level, interest groups pressure the government to adopt favorable policies. Internationally, the government seeks to fulfill domestic preferences “while minimizing the adverse consequences of foreign developments” [@putnam1988 p. 434]. While the theory has been expanded to cover multi-levels games [see @collinson1999], its core remains the same: the outcome of negotiations should lie within the overlap of diverse interest groups. Consequently, presidential speeches are not necessarily contradictory across venues but a common denominator of the interests of different groups. Adopting two-level games to explain American presidential speeches, @calderwood2019 and @calderwood2020 find supporting evidence of this theory.

The two-level game theory implies that actors respond rationally to systemic pressures imposed by their position in relation to interest groups. This view seems incompatible with the possibility of contradictory policies and of varied problem-constructions depending on specific historical contexts, timing, and places. Rather, we start from the idea of transnationalism as relationships that transcends nation-states and incorporates non-state actors, spamming from the local level to the international level [@nye1971; @keck1998]. In this context, international summits are considered a transnational setting because non-state actors participate and shape them. We view, for instance, summits as the United Nations Framework Convention of Climate Change (UNFCCC) not only as a purely intergovernmental body, but as a transnational space composed of international media, corporations, non-governmental organizations, and social movements.

@keck1998 theory of international politics as transnational networks rejects the dichotomy of international and domestic levels arguing that the identity and goals of state and non-state actors are not derived from their structural location vis-a-vis domestic and foreign interest groups. Rather, international politics as transnational networks have “a structured and structuring dimension” [@keck1998, p. 4]: state and non-state actors participate in and shape international politics.

While their theory is developed in the context of value-based advocacy groups, hence non-state actors, the overall view of international politics as structured and structuring, should hold for state actors as well. Important participants of transnational networks, as presidents, can use myriad of tactics to introduce, define, persuade, socialize, and pressure when interacting with other actors. They can create politically meaningful information and/or move it where they believe it can have an impact [@keck1998, p. 24]. They can also leverage institutions to realize symbolic or material gains. In this sense, the positions of actors in transnational networks derive from complex interactions in which their identities and goals are malleable and can vary depending on the space and time of the interaction.

Problem-constructions are choices made by presidents when speaking. If malleable identities across space and time are a defining feature of transnational politics, it follows that presidents' choice of problem-constructions depend on the president's views of an audience at a certain period. Specifically, how urgent a problem is in a certain space and/or at a certain time.

@@

Presidents construct problems and attempt to inform and/or leverage transnational audiences for symbolic or material gains according to their expectations of that audience. We conceptualize spaces in terms of settings. Settings are levels of aggregation for transnational audiences that make sense in light of the object of policy under study. This indicates that how presidents construct policy objects as problems can vary, for instance, within the country — speaking inside the Amazon region versus outside — because presidents identify these audiences as sources of distinct types of support and demand. The same holds true when speaking abroad — in Amazonian countries, non-Amazonian countries, and at international summits — where presidents’ expectations and identities of an audiences could diverge completely from that of domestic actors. In each one of these settings different problems might be more urgent and the president's expectations about what can be leveraged for material and symbolic gains might influence how a problem is implicitly or explicitly constructed.

@@

How presidents construct policy objects can also vary over time because the urgency of certain problems changes as new information arises. We conceptualize time as the perceived urgency of a broad and related issue to those surrounding a policy object. This indicates, for instance, that as climate change becomes more urgent and mainstreamed as a broad issue, it is more likely that actors will consider it when discussing various forms of public policy. It is likely that when speaking about subsidies for fossil fuel companies in the 1980s, presidents would not mention the dangers of climate change, whilst today, climate change would likely be mentioned given the increased global attention.

While time and space might interact, here we are interested in their cumulative effects: when speaking in the 1990s to an audience that finds climate change urgent, certain transnational actors would likely mention it as a problem independently of the perceived urgency of a climate breakdown at the time. Accordingly, certain transnational actors might even consider touching upon the issue while speaking at a business association today, because of its urgency.

Our conceptual framework, thus, combines Bacchi’s theory of problem-representation and Keck and Sikkink’s theory of transnationalism to account for diverse problem-constructions. We help to advance the literature about the Amazon in two ways. First, we provide an outlook of the region, its people and forest beyond its environmental issues by focusing on different problem-constructions. In doing so, we avoid assuming what type of problem the Amazon is and, instead, focus on how it has been represented as a problem. Second, many analyses of Brazilian environmental policies dissect its national and international nature but do not adopt methodological strategies that distinguish how important actors behave differently in both scenarios [@hochstetler2007]. Our focus on presidential speeches affords us analytical leverage to identify transnational variation, as speeches take place in multiple settings over time. Finally, presidential speeches in Brazil have been studied for topics such as race relations [@silva2019] and inequality [@grangeia2017]. When it comes to the environment, studies tend to focus primarily on Bolsonaro’s discourse [@marquardt2022; @mendes2022] or policies [see @capelari2023]. We argue, however, that a longitudinal view across Brazilian presidents is important to understand how much continuity and change there really is.

@@

To create a dataset containing all Brazilian presidential speeches,we update a dataset of all Brazilian presidential speeches from 1985 to 2019 [@cezar2020] to include speeches from 2020 to 2022.The final dataset contains 6,240 speeches with their date and location. We proceed to identify speeches that refer to the Amazon as a region, people, or forest, by detecting the stem “amazon”. In Portuguese, the stem captures terms such as “Amazonia”, “Amazonica”, “Amazonidas”, “Amazonense(s)”, “Amazonas”, among others. We find that 968 speeches are, at least partially, about the Amazon.

@@

*National Sovereignty* 
In the process of securing Amazonian borders since the 18th century, Brazil thwarted “the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru” [@hecht1990, p. 8], and secured the Amazon as Brazilian. While military diplomacy was very successful, the process did not come without its issues and sovereignty in the Amazon became one of the main military concerns[@rajao2020]. More recently, as we move to a world where non-state actors gain importance in international politics [@andonova2014; @keck1998], sovereignty-related problems become more varied. Threats to national sovereignty, consequently, broaden from nation-states to a wide set of actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil and any foreign or non-state presence in the region can be part of a broad strategy to take the region. The policy solutions to the issue of sovereignty included the monitoring of the borders and strict regimes related to entry into the region.

*Economic Integration* 
The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon [@becker2005]. In 1966, the Brazilian Military launched ‘Operation Amazon’, a policy to modernize the region based on three assumptions [@acker2021]. First, it assumed that nature should be conquered by mankind. Second, it assumed that exploiting natural resources would render the Amazon region economically profitable. Third, it assumed that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructure projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region into the national and international economy.

*Environmental Conservation*
The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of domestic environmental institutions [@drummond2006]. The creation of these institutions stemmed from the impression of the lack of control over the environment following the efforts to economically integrate in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987]. The environmental conservation problem-construction emphasizes that the Amazon should be preserved by the creation of protected areas and deforestation should be halted. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

*Social Development*
As democracy established itself in Brazil, especially with the 1988 constitution, settlers in the Amazon became electorates with constitutional rights. Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples’ dignity, standards of living, and other constitutional rights. However, such emphasis can be masked within both the economic integration and environmental conservation accounts. While for some groups, such as indigenous peoples, constitutional rights and preservation of the forest go hand in hand, providing better standards of living to large portions of populations of the Amazonian states and conservation are not necessarily connected. Social development is a problem-construction that focuses on constitutional rights, citizenship, and dignity. The policy solution implies enabling access to water, sanitation, electricity, internet, and radio, as well as the construction of schools and hospitals.

@@

Using a random sample of 1024 Amazonian statements and the codebook (see appendix), we hand-coded 50% of our data to train a machine learning model, yielding an average intercoder agreement of 85%. Any disagreements in coding were resolved through discussion and refinement of the codebook. We chose to hand code half of the observations because there are several nuances in how presidents talk about the Amazon over time, and thus the training set size should be increased [see @grimmer2022].

We then randomly divided the hand-coded data into a training set (80% of observations) and validation set (20% of observations). We use a support-vector machine algorithm, a non-probabilistic linear classifier appropriate for binary categories[@meyer2021; @noble2006], to train a model to label text. The model was validated by testing it on the validation set and it was 95% accurate in coding these statements. The final dataset, excluding false positive matches, contains 1924 coded Amazonian statements. ^[We also train the model to identify false positive matches in Amazonian statements. These are statements that contain the stem “amazon” but are not meaningful as they might salute the governor of the Amazon, for example.]

>>>>

---
title: One Amazon Problem? Problem-constructions and Transnationalism in Brazilian Presidential Discourse since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Sociology and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences and affiliated to the Centre for International Environmental Studies.]
date: 17 February 2023
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\textbf{Word Count:}#1}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: The Amazon is a complex object of policy as it comprises environmental, economic, social, and sovereignty concerns. Despite this complexity, governments are often portrayed as having a single understanding of the Amazon as a political problem. We investigate how the Amazon has been constructed as a problem in Brazilian presidential speeches. To do so, we develop a framework that accounts for how important actors in transnational politics, as presidents, construct policy objects as particular problems depending on where and when they speak and when. Using supervised machine learning, we classify statements about the Amazon in 6240 presidential speeches since 1985. We find that presidents often construct the Amazon as an environmental problem when speaking far away from the Amazon, whereas they usually construct it as economic or social problems when in the Amazon. Furthermore, presidents increasingly construct the Amazon as multiple problems rather than only economic or environmental problems.
---

```{=latex}
\keywords{ discourse analysis, presidential speeches, transnational governance, environmental policy, Brazilian Amazon, supervised learning}
```
```{=latex}
\wordcount{ 8190 (including abstract, text, references, and footnotes)}
```
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(ggplot2)
library(plm)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

# Introduction

The Brazilian Amazon, which covers around 60% of the whole biome, is home to over 28 million people. The Rainforest spans over seven international borders and contains enormous amounts of highly valued minerals such as iron ore and gold. Its most inhabited city, Manaus, enjoys a distinct fiscal regime designed to create jobs in the region and connect its economy to the rest of the country. At the same time, an extensive portion of the Brazilian Amazon is protected by a combination of indigenous territories and protected areas. The Brazilian Amazon is, thus, a complex object of policy as it comprises environmental, economic, social, and sovereignty concerns. Despite this complexity, governments are often portrayed in the literature about policies in the Amazon as having a single understanding of what the Amazon problem is and consequently, proposing coherent policies for the region. Although it is helpful to compare variation across different governments, this analysis oversimplifies policies and discourse within governments misrepresenting what, in reality, is more intricate and inconsistent. In this article, we ask how has the Amazon been constructed as a political problem since 1985?

We combine @bacchi1999's theory of problem representation and @keck1998's theory of transnationalism to develop a framework outlining how transnational actors, as Presidents, construct objects of policies as particular problems depending on where and when they participate in politics. To examine this variation, we created a dataset containing 6240 official speeches by all Brazilian presidents since 1985. Drawing from the socio-cultural history of the Brazilian Amazon, we inductively developed a codebook to categorize problems related to political concerns over the Amazon. These problem-constructions relate to national sovereignty, economic integration, social development, or environmental conservation. We employ supervised machine learning to classify how the Amazon related statements construct it as a problem, and then conduct an inferential and descriptive analysis of the data. A systematic focus on presidential speeches is important because while many scholars recognize decisions to deforest on the ground are derived from what the government says and does [@capobianco2021; @campbell2015], we lack empirical accounts of the Amazon in presidential discourse.

We have four main findings. First, when presidents are far away from the Amazon region, they are more likely to construct the Amazon as an issue of environmental conservation. Within the Amazonian region, presidents usually construct the Amazon as an issue of social development and economic integration. Second, presidents are more likely to construct the Amazon as an environmental conservation problem when environmental related issues are not urgent (e.g. lowering deforestation rates). When environmental related issues become more urgent (e.g. rising deforestation rates), presidents tend to side-step the issue and construct the Amazon as an issue of economic integration. Third, we find that high-profile environmental events, which receive strong international media attention, drive up the frequency at which the Amazon appears in discourses, but does not systematically change how the Amazon is constructed as a problem. Finally, we find that presidents are increasingly more likely to mix problem-constructions and construct the Amazon as a complex and multifaceted object of policy.

Empirically, we contribute to the literature about the Amazon by providing the first overview of where, how, and when the Amazon has been constructed as a problem in presidential speeches. We show how the same government constructs the Amazon as a different problem depending on where they speak and how urgent an aligned issue is. Analyses of Brazilian environmental discourse and policies do not adopt methodological strategies that distinguish how important actors behave differently inside and outside the country, by doing so we reveal how the Amazon as a problem varies in transnational politics. Conceptually, problem-constructions highlight the extent to which Amazonian problems are ignored or privileged in transnational politics, and avoids assumptions of the Amazon as a solely environmental problem.

This article is divided into 4 sections. First, we outline the connection between problem-constructions, transnationalism, and presidential speeches. Next, we present our research design including the description of our data and methodological approaches. In the third part, dedicated to the analysis, we display the results of our inferential models and move to explain developments in problem-constructions over time. We conclude by summarizing the empirical and conceptual contributions of the article.

@@

Scholars writing about policies in the Brazilian Amazon often describe federal governments as proponents of a cohesive set of policies toward the Amazon [see @drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. The 1964 military dictatorship, for example, is associated with securing sovereignty in the region by populating it and integrating it into the national economy [@drummond2006]. Whereas, governments from the late-1980s up to the late-2000s are associated with a turn towards policies that focus on environmental conservation of the Amazon [@padua2012]. The presidencies of Rousseff (2011-2016) and Temer (2016-2018) are connected to the de-prioritizing of environmental conservation policies, while Bolsonaro (2019-2022) with the dismantling of environmental policies [@capobianco2019]. These works are important to understand how different governments acted towards the Amazon, but they represent governments monolithically by bundling them into cohesive policy cycles.

Elsewhere, social scientists have shown that there is more variation and contradiction within a specific government's policies and discourse [see @lopez2023; @fairfield2017; @poncedeleon2021] than suggested by the literature above. We argue that the reason behind monolithic representations of governments in literature about the Amazon is conceptual: many studies of the region contain an explicit or implicit assumption of what the "Amazon problem" is. Because of the Amazon Rainforest role in global climate-mitigation, , and consequently, as an environmental problem. If the Amazon is thought of as an environmental problem, then it demands an environmental solution. The Amazon region, its peoples, and forests, though, are a historically complex object of policies that comprise economic, social, and sovereignty concerns.

@bacchi1999 [p. 63] places the question of "what is the problem represented to be" in the center of policy studies. The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed. Hence, discourse represents, implicitly or explicitly, objects of policies as instances of a specific problem.Problem-representation matters because it carries implications related to who is involved both in the root causes and in the solutions of the problem [@bacchi1999]. It also helps us shed light on what problems are ignored or privileged.

@@

Presidential discourses are a good instance to identify problem-constructions over time. Presidential speeches and policy share a co-constitutive relationship as speeches have the power to introduce, define, and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004]. This is particularly true in the Brazilian context where the executive branch of government, led by the president, has historically been the primary player in shaping public policy, especially after the 1988 constitution [@macaulay2017]. In Brazil's hybrid system, often referred to as a Coalitional Presidentialism [@couto2021], presidents build legislative support for a broad agenda through the distribution of positions in the state bureaucracies. This means that presidents often have enough power and support to implement the agenda they deem appropriate. We thus consider presidential speeches an adequate instance to identify various problem-constructions for the same policy object.

@@

The possibility of different problem-constructions for the same policy object in political discourse requires an explanation of why they vary. There are two potential theories for explaining variation in problem-construction in international/transnational politics: Putnam's two-level game and Keck and Sikkink's transnational networks.

@putnam1988 argues that the outcomes of international negotiations lie within the overlap between the interests of domestic and international groups. At the national level, interest groups pressure the government to adopt favorable policies. Internationally, the government seeks to fulfill domestic preferences "while minimizing the adverse consequences of foreign developments" [@putnam1988 p. 434]. The theory has been expanded to cover multi-levels games [@collinson1999], but its core remains the same: the outcome of negotiations should lie within the overlap of distinct interest groups. Consequently, presidential speeches are not necessarily contradictory across venues but a common denominator of the interests of different groups. Adopting two-level games to explain American presidential speeches about climate change, @calderwood2020 found supporting evidence that location does not change the content of the speech, but can lead presidents to side-step the issue.

The two-level game theory implies that actors respond rationally to systemic pressures imposed by their position in relation to interest groups. This view seems incompatible with the possibility of contradictory policies and of varied problem-constructions. Rather, we start from the idea of transnationalism as relationships that transcends nation-states and incorporates non-state actors, from the local level to the international level [@nye1971; @keck1998]. In this context, international summits are considered a transnational setting because non-state actors participate in and shape them. We view, for instance, summits as the United Nations Framework Convention of Climate Change (UNFCCC) not only as a purely intergovernmental body, but as a transnational space composed of international media, corporations, non-governmental organizations, and social movements.

While @keck1998 theory is developed in the context of value-based advocacy groups, hence non-state actors, the overall view of international politics holds for state actors. Their theory rejects the dichotomy of international and domestic levels and argues that the identity and goals of transnational actors are not derived from their structural location vis-a-vis domestic and foreign interest groups. Rather, international politics as transnational networks have "a structured and structuring dimension" [@keck1998, p. 4]: state and non-state actors participate in and shape international politics.

Important participants of transnational networks, as presidents, can use a myriad of tactics to introduce, define, persuade, socialize, and pressure when interacting with other actors. They can create politically meaningful information and/or move it where they believe it can have an impact [@keck1998, p. 24]. They can also leverage institutions to realize symbolic or material gains. In this sense, the positions of actors in transnational networks derive from complex interactions in which their identities and goals are malleable and can vary depending on the space and time of the interaction.

Problem-constructions are choices transnational actors have of how to represent objects of policies as problems. If malleable identities across space and time are a defining feature of transnational politics, it follows that transnational actors', as presidents, choice of problem-constructions depend on their views of an audience at a certain period. Specifically, how urgent a problem is in a certain space and/or at a certain time.

@@

Transnational actors construct problems and attempt to inform and/or leverage audiences for symbolic or material gains according to their expectations of that audience. We conceptualize space in terms of settings. Settings are levels of aggregation for transnational audiences that make sense in light of the object of policy under study. This indicates how transnational actors construct policy objects as problems can vary, for instance, within the country --- speaking inside the Amazon region versus outside --- because presidents identify these audiences as distinct. The same holds true when speaking abroad --- in Amazonian countries, non-Amazonian countries, and at international summits --- where presidents' expectations and identities of an audience could diverge completely from that of domestic actors. In each one of these settings different problems might be more urgent and the president's expectations about what can be leveraged for material and symbolic gains might influence how a problem is implicitly or explicitly constructed.

@@

How presidents construct policy objects varies over time as the urgency of certain problems changes not only across settings but in time as new information arises. We conceptualize time as the urgency of a broad and related issues to those surrounding a policy object. This is largely defined by factors exogenous to the speaker. This indicates, for instance, that as climate change becomes more urgent and mainstreamed as a broad issue, it is more likely that actors will consider mentioning climate change when discussing various forms of public policy. For instance, when speaking about subsidies for fossil fuel companies in the 1980s, presidents would likely not mention the dangers of climate change, whilst today, climate change would likely be mentioned given the increased relevance of the topic.

While time and space interact, here we are interested in their cumulative effects: when speaking in the 1990s to an audience that finds climate change urgent, certain transnational actors would likely mention it as a problem independently of the perceived urgency of a climate breakdown at the time. Accordingly, certain transnational actors might even consider touching upon the issue while speaking at a business association today, because of its urgency.

Our conceptual framework, thus, combines Bacchi's theory of problem-representation and Keck and Sikkink's theory of transnationalism to account for diverse problem-constructions. We help to advance the literature about the Amazon in three main ways. First, we provide an outlook of the region, its people, and forest beyond its environmental nature by focusing on different problem-constructions. In doing so, we take a step back and avoid assuming what type of problem the Amazon is. Instead, we focus on how the Amazon has been represented as a problem. Second, many analyses of Brazilian environmental policies center on its national and international nature without adopting methodological strategies that distinguish how important actors behave differently in both scenarios [@hochstetler2007]. Our focus on presidential speeches afford us the analytical leverage to identify transnational variation, as speeches take place in multiple settings over time. Finally, presidential speeches in Brazil have been studied for topics such as race relations [@silva2019] and inequality [@grangeia2017]. When it comes to the environment, studies tend to focus primarily on Bolsonaro's discourse [@marquardt2022; @mendes2022]. We argue, however, that a longitudinal view across Brazilian presidents is important to understand how much continuity and change there is.

If our conceptual framework holds, it entails that transnational actors view different settings as distinct policymaking instances. While there might exist convergences of priorities over time, as problems become more urgent, contradictions become possible. This does not mean that policies and priorities change instantaneously and consequently do not matter. Rather, presidents will focus on representing issues in forms aligned with what their perception of an audience are.

@@

To create a dataset containing all Brazilian presidential speeches, we update a dataset of all Brazilian presidential speeches from 1985 to 2019 [@cezar2020] to include speeches from 2020 to 2022. The final dataset contains 6240 speeches with their date and state or country location. We proceed to identify speeches that refer to the Amazon as a region, people, or forest, by detecting the stem "amazon". In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 968 speeches are, at least partially, about the Amazon.

@@

*National Sovereignty* In the process of securing Amazonian borders since the 18th century, Brazil thwarted "the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru" [@hecht1990, p. 8], and secured the Amazon as Brazilian. While military diplomacy was very successful, the process did not come without its issues and sovereignty in the Amazon became one of the main military concerns[@rajao2020]. More recently, as we move to a world where non-state actors gain importance in international politics [@andonova2014; @keck1998], sovereignty-related problems become more varied. Threats to national sovereignty, consequently, broaden from nation-states to a wide set of actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil and any foreign or non-state presence in the region can be part of a broad strategy to take the region. The policy solutions to the issue of sovereignty included the monitoring of the borders and strict regimes related to entry into the region.

*Economic Integration* The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon [@becker2005]. In 1966, the Brazilian Military launched 'Operation Amazon', a policy to modernize the region based on three assumptions [@acker2021]. First, it assumed that nature should be conquered by mankind. Second, it assumed that exploiting natural resources would render the Amazon region economically profitable. Third, it assumed that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructure projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure to integrate the region into the national and international economy.

*Environmental Conservation* The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of domestic environmental institutions [@drummond2006]. The creation of these institutions stemmed from the impression of the lack of control over the environment, following the efforts to economically integrate in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987]. The environmental conservation problem-construction emphasizes that the Amazon should be preserved by the creation of protected areas and deforestation should be halted. The policy solutions imply, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

*Social Development* As democracy established itself in Brazil, especially with the 1988 constitution, settlers in the Amazon became electorates with constitutional rights. Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples' dignity, standards of living, and other constitutional rights. However, such emphasis can be masked within both the economic integration and environmental conservation accounts. While for some groups, such as indigenous peoples, constitutional rights and preservation of the forest go hand in hand, providing better standards of living to large portions of populations of the Amazonian states and conservation are not necessarily connected. Social development is a problem-construction that focuses on constitutional rights, citizenship, and dignity. The policy solutions imply enabling access to water, sanitation, electricity, internet, and radio, as well as the construction of schools and hospitals.

@@

We randomly select a sample of 1024 Amazonian statements (50% of our data) and each author hand-codes it separately. Once manual coding was finished, the average intercoder agreement was 86%, the authors resolved disagreements through discussion and refined the codebook (see appendix). We then randomly divided the hand-coded data into a training set (80% of observations) and validation set (20% of observations) to train a machine learning algorithm. We chose to hand code half of the observations because there are several nuances in how presidents talk about the Amazon over time, and thus the training set size should be increased [see @grimmer2022].

We use a support-vector machine algorithm, a non-probabilistic linear classifier appropriate for binary categories [@meyer2021; @noble2006], to train a model to label text. The model was validated by testing it on the validation set. As a robustness check, the authors randomly select 100 automatically coded statements and find that the automated model is 95% accurate in coding these statements as the authors would. The final dataset, excluding false positive matches, contains 1924 coded Amazonian statements [^1]. Automating the code saved the authors about two months.
